---
title: "Making Box Plots for AACR Poster"
author: Jaron Magstadt
date: August 2024
output: html_notebook
---

## Notebook Tasks

The things I did in this notebook are:

- Read in NTA data from canine exosome samples
- Filtered data to only include relevant conditions/results
- Generated summary statistics using the `summarize` function
- Created box plots displaying data of interest
- Made the box plots look **pretty!**

## Setup Steps

To setup, I started by loading in the packages I needed.

```{r library}
library(tidyverse)
```
I continued setup by setting my working directory and paths.

```{r paths}
#I assigned the path to my working directory to a variable and then set my wd.
main_dir <- file.path("~","Documents","repos","ALSF")
setwd(main_dir)

#I assigned the path for my data directory to a variable.
data_dir <- file.path(main_dir, "data")

#I set up a directory for my plots to go to
plots_dir <- file.path(main_dir,"plots")
fs::dir_create(plots_dir)
```

## Reading in Data

Next, I read in my data of interest. The data of interest is a data frame containing NTA data from several canine exosome samples. The data was previously saved in a `.csv` format, so I used the `read.csv` function to read in my data.

```{r data read}
NTA_data <- read_csv(file.path(data_dir, "NTA_concentrations_cleaner.csv"))
```
It worked! **YAY!**
## Viewing Data

Like Stephanie says, **always** look at your data! My data frame isn't terribly large, but still unwieldy, so I will use the `head()`, `summary()`, and `names()` functions to look at my data.

```{r data peek}
head(NTA_data)
summary(NTA_data)
names(NTA_data)
```
That's a lot of information! But we don't need all of ithe information in the data frame to do the analyses that I am interested in, so lets do some filtering.

## Filtering and Summarizing Data

There are many types of cancer/controls in the data frame I read into R. I am only interested in `"OSA"` and `"HC"`, so I created a character vector that contained these labels to help with an upcoming step.

```{r groups}
my_groups <- c("OSA","HC")
my_groups
```
Now that I had a vector containing the names of the conditions I'm interest in, I was able to create a pipeline to create a summary of the average size and concentration of exosomes in the `"OSA"` and `"HC"` conditions. I used the `filter()` function along with the `mygroups` vector to only keep the conditions I was interested in, then used the `summarize` function to make a table of summary statistics.

```{r filtering and summarize}
NTA_summary <- NTA_data |> 
  filter(Group %in% my_groups) |> 
  group_by(Group) |> 
  summarize(Mean_Size = mean(Size),
            Std_Dev_Size = sd(Size),
            Mean_Concentration = mean(Concentration),
            Std_Dev_Concentration = sd(Concentration)
  )
 NTA_summary
```
Look at that! I can now see the mean size and concentration, as well as standard deviations, for the `"HC"` and `"OSA"` conditions!

## Making Data Frames to Use With ggplot2

Now it was time time move onto the main objective: **making pretty plots for a poster.**

Well...**almost**...

In order to make plots, I wanted to have more concise data frames that only had the data I was interested in. This wasn't entirely necessary, but I wanted to make sure the data was pulling into the plot correctly, and doing things this way ensured that, at least in my mind. I did this by using the `filter()` function again in a set of two pipelines: one for exosome size and one for exosome concentration.
```{r more filtering}
 #Making size table for plot
 NTA_size <- NTA_data |> 
   filter(Group %in% my_groups) |> 
   group_by(Group) |> 
   select(Size)

 NTA_size   
```

```{r more filtering 2}
 #Making concentration table for plot
 NTA_conc <- NTA_data |> 
   filter(Group %in% my_groups) |> 
   group_by(Group) |> 
   select(Concentration)
 
NTA_conc   
```
## Making Box Plots using ggplot2


```{r size box plot}
#Graphing Size
 NTA_size_bp <- ggplot(
   NTA_size,
   aes(
     x = Group,
     y = Size,
     fill = Group,
     )
 )+ geom_boxplot(lwd = 1, fatten=1) +
   theme_bw() +
   theme(plot.title = element_text(size = 14, face = "bold", hjust=0.5)) +
   scale_fill_manual(values=c("#8a0a25", "#f7ba4d")) +
   labs(
     x = "Condition",
     y = "Size",
     title = "Exosome Size in Healthy Control vs. Osteosarcoma"
   )
 
 NTA_size_bp
```


